# IV-Strategy 


## Data {.unnumbered}

First, we import the dataset called “INEGI employment.dta”. This file corresponds to a panel dataset of Mexican commuting zones (CZ), which contains the following variables: 

- **year:** Year of the Economic Census. 

- **CZ:** Commuting zone (the union of several municipalities). 

- **country code:** The country of ownership of establishments employing workers in a given CZ and year. 

- **workers:** The total number of employees working for an establishment whose ownership comes from country code in a given CZ and year. 

```{r}
#| echo: true
#| message: true
#| warning: false
#| output: true
#| 

# Import libraries
library(haven)
library(kableExtra)

# Import the data
INEGI <- read_dta("./exercise_1/INEGI_employment.dta")

# Glimpse the data
head(INEGI, 5) %>% 
  kable() %>%
  kable_styling(position = "center")

```


## Question 1: Create Variables {.unnumbered}


We must create diverse variables, the outcome of interest, the explanatory variable, and the instruments. For this example, I choose the impaired numbers $1,3,5$ for the instruments.

- Outcome: $\Delta \ell_{c z, t}=\log \left(L_{c z, t}^D\right)-\log \left(L_{c z, t-5}^D\right)$

- Explanatory:  $\hat{X}_{c z, t} \equiv \frac{L_{c z, t}^F-L_{c z, t-5}^F}{L_{c z, t-5}}$

- Instrument 1: $\hat{Z}_{c z, t}^1 \equiv \sum_o \frac{\sum_{c z^{\prime} \neq c z}\left(L_{c z^{\prime}, t}^{F_o}-L_{c z^{\prime}, t-5}^{F_o}\right)}{\sum_{c z^{\prime}} L_{c z^{\prime}, 1994}} S_{c z, 1994}^o$

- Instrument 2: $\hat{Z}_{c z, t}^3 \equiv \sum_o \frac{\sum_{c z^{\prime} \neq c z}\left(L_{c z^{\prime}, t}^{F_o}-L_{c z^{\prime}, t-5}^{F_o}\right)}{\sum_{c z^{\prime}} L_{c z^{\prime}, t-5}^{F_o}} S_{c z, t-5}^o$

- Instrument 3: $\hat{Z}_{c z, t}^5 \equiv \sum_o \frac{\sum_{c z^{\prime} \neq c z}\left(L_{c z^{\prime}, t}^{F_o}-L_{c z^{\prime}, t-5}\right)}{\sum_{c z^{\prime}} L_{c z^{\prime}, t-5}^{F_o}} S_{c z, 1994}^o$


### Outcome and Explanatory {.unnumbered}



First, I would like to simplify the data set; we are only interested in the domestic (Mexico) and foreign (everything else) markets. So I'm collapsing the workers into domestic and foreign groups.


```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# Import libraries
library(tidyverse)

# Collapse workers in domestic and foreign markets
data_ox <- INEGI %>% 
  mutate(domestic = ifelse( country_code == "MEX", "D", "F")) %>% 
  group_by(year,CZ,domestic) %>% 
  summarize( workers = sum(workers)) %>% 
  ungroup()

# Glimpse the data set
head(data_ox , 5) %>% 
  kable() %>%
  kable_styling(position = "center")

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| output: true

# Glimpse the data set
head(data_ox , 5) %>% 
  kable() %>%
  kable_styling(position = "center")

```

With the simplified data set, we now add two columns: 

1. The sum of the total labor in period $t$.
2. A lagged column with the domestic and foreign labor of $t-5$ periods.


```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# 1) Sum of labor
  
  # a) Aux data set with the total labor per CZ and year
  total <- data_ox  %>% 
    group_by(year,CZ) %>% 
    summarize( workers_total = sum(workers)) %>% 
    ungroup()
  
  # b) Join aux data set with domestic data
  data_ox  <- data_ox  %>% 
    left_join(total, by = c("year" ,"CZ"))
  


# 2) Lagged column t-5 of domestic, foreign, and total workers
  
  data_ox  <- data_ox  %>% 
    group_by(CZ,domestic) %>% 
    mutate(
      workers_lag = lag(workers, 1),
      workers_total_lag = lag(workers_total, 1)
    ) %>% 
    arrange(CZ) %>% 
    ungroup()
  
# 3) Glimpse the data set
  
head(data_ox , 5) %>% 
  kable() %>%
  kable_styling(position = "center")

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| output: true

# Glimpse the data set
head(data_ox , 5) %>% 
  kable() %>%
  kable_styling(position = "center")

```

We can now calculate the outcome and explanatory variable

```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# 1) Outcome variable 
    outcome_data <- data_ox  %>% 
      mutate(
        outcome = ifelse( domestic == "D", log(workers) - log(workers_lag), NA )
      ) %>% 
      filter( !is.na(outcome))%>% 
    select(year, CZ, outcome)


# 2) Explanatory
    explan_data <- data_ox  %>% 
      # Create the variable
      mutate(
        explanatory = ifelse( domestic == "F", (workers - workers_lag) / workers_total_lag, NA )
      ) %>% 
      # Remove NAs
      filter( !is.na(explanatory))%>% 
      # Keep variables of interest
    select(year, CZ, explanatory)

```

### Instrumental variables {.unnumbered}

We now pass to the instrumental variables; I had difficulty understanding your notation, so my thought process might not be as clear. I will try to be as specific as possible.


1. Every instrument for each $o$ has $\sum_{c z^{\prime} \neq c z}\left(L_{c z^{\prime}, t}^{F_o}-L_{c z^{\prime}, t-5}^{F_o}\right)$; let's try to calculate this first.

```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# 1) Get the difference in labour for each cz, o, t

  data_iv <- INEGI %>% 
    # Remove domestic market
    filter( country_code != "MEX") %>% 
    group_by(CZ,country_code) %>% 
    mutate(
      # Get the lag for each cz and o 
      workers_lag_o = lag(workers, 1), 
      # Calculate the differnce
      diff_workers = workers - workers_lag_o
    ) %>% 
    ungroup() 

# 2) Calculate the sum of very cz' != cz
  
  # a) Calculate the sum of every cz
  aux <- data_iv %>%
    filter( !is.na(workers_lag_o) ) %>% 
    group_by( year, country_code) %>% 
    summarize(sum_cz = sum(diff_workers))
  
  # b) Add aux to data_iv dataset and substract the cz at each row
  
  data_iv <- data_iv %>% 
    left_join(aux, by = c("year", "country_code")) %>% 
    mutate( sum_cz_prime = sum_cz - diff_workers)
    
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| output: true

# Glimpse the data set
head(data_iv , 10) %>% 
  kable() %>%
  kable_styling(position = "center")

```

2. Instruments 3 and 5 share the following term for every $o$ has $\sum_{c z^{\prime}} L_{c z^{\prime}, t-5}^{F_o}$; I'll focus on these two instruments due to lack of time.

```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# 1) For every o and t we must sum all the cz's lagged labor 
  
  # a) Calculate the sum
  aux <- data_iv %>%
    ungroup() %>% 
    filter(!is.na(workers_lag_o ) ) %>% 
    group_by(year, country_code) %>% 
    summarize( sum_lag = sum(workers_lag_o))

  
  # b) Add aux to data_iv dataset 
  data_iv <- data_iv %>% 
    left_join(aux, by = c("year", "country_code")) 
    
```


3. We can now calculate $\alpha_o = \frac{\sum_{c z^{\prime} \neq c z}\left(L_{c z^{\prime}, t}^{F_o}-L_{c z^{\prime}, t-5}^{F_o}\right)}{\sum_{c z^{\prime}} L_{c z^{\prime}, t-5}^{F_o}}$ for instruments 3 and 5.

```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# Calculate alpha_o and remove auxiliar variables
  data_iv <- data_iv %>%
    mutate( alpha_o = sum_cz_prime / sum_lag) %>% 
    select(-sum_cz,-sum_cz_prime, -sum_lag)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| output: true

# Glimpse the data set
head(data_iv , 10) %>% 
  kable() %>%
  kable_styling(position = "center")



```

4. We calculate both instruments 3, and 5. Also, we integrate the final data set for the analysis.

```{r}
#| echo: true
#| message: false
#| warning: false
#| output: false

# 1) Calculate the share S^o_cz t-5

  # a) Auxiliar data set with total workers from t-5
  total <- total %>% 
    group_by(CZ) %>% 
    mutate( workers_total_lag = lag(workers_total, 1) ) %>% 
    ungroup() %>% 
    arrange(CZ)

  # b) Share of t-5
  data_iv <- data_iv %>% 
    left_join(total, by = c("year","CZ")) %>% 
    mutate(share_o = workers_lag_o / workers_total_lag)
  
# 2) Calculate the share S^o_cz 1994
  
  # a) Get the share 1994 of every cz and o
  aux <- data_iv %>% 
    filter( year == 1994) %>% 
    mutate(share_94 = workers / workers_total) %>% 
    select( CZ, country_code, share_94)
  
  # b) Add the variable to the data_iv table
  
  data_iv <- data_iv %>% 
    left_join(aux, by = c("CZ","country_code")) %>% 
    select(year, CZ, country_code, alpha_o, share_o, share_94)
  
# 3) Multiply alpha_o * share_o /94 and calculate final instruments
  
   aux <- data_iv %>% 
    filter(!is.na(alpha_o)) %>% 
    mutate(
      z_3 = alpha_o * share_o,
      z_5 = alpha_o * share_94
    ) %>% 
    group_by(year, CZ) %>% 
    summarize(
      z_3 = sum(z_3),
      z_5 = sum(z_5)
    )
  
# 4) integrate the final data set
  
  model <- aux %>% 
    left_join(outcome_data, by = c("year", "CZ")) %>% 
    left_join(explan_data, by = c("year", "CZ")) 
  
  
    
```


```{r}
#| echo: false
#| message: false
#| warning: false
#| output: true

# Glimpse the data set
head( filter(model, !is.na(z_3)) , 5) %>% 
  kable() %>%
  kable_styling(position = "center")

rm(aux, data_iv, data_ox, explan_data, outcome_data, total)

```


## Question 2 {.unnumbered}

Present a summary statistics table for these five variables by Census. Include the $1^{\text {st }}$ and $99^{\text {th }}$ percentiles, median, mean, and standard deviation for each variable.


```{r}
#| echo: true
#| message: false
#| warning: false
#| output: true

# Summary statistics

 q2 <- model %>% 
  ungroup() %>% 
  select( -year, -CZ) 

 q2 %>% 
    summary() %>% 
    kable() %>%
    kable_styling(position = "center")


```

## Question 3 {.unnumbered}

::: {.callout-note}
We want to implement an IV strategy to identify the causal effect of $\hat{X}_{c z, t}$ on $\Delta \ell_{c z, t}$. $\hat{Z}_{c z, t}^i$ for $i \in\{1, \ldots, 6\}$ are candidates for an IV. Which of them could be written with a shift-share structure? Write the number of the instruments that correspond to the shift-share structure and complete the algebra to obtain the $\Delta L_{c z, t}^{F_o}$ of the instruments that follow the shift-share structure.
:::

**Answer:** I don't have enough time for algebra, so I'll try a deductive approach. 


As stated by [Ferri, 2022](http://fmwww.bc.edu/EC-P/wp1053.pdf), the essence of the Shift-Share approach is to decompose the endogenous explanatory variable as an accounting identity with multiple component parts; preserve that which is most exogenous in the accounting identity, and neutralize that which is most endogenous. 

Endogeneity is neutralized via delocalizations over space and time. That is, the more endogenous component - the Shift vector - is replaced with nonlocal averages (a strong delocalization); and the more exogenous component - the Share vector - is lagged (a weaker delocalization).


Ferri frames the structure of a shift-share instrument:
$$
\tilde{\mathrm{X}}_{\mathrm{t}, \mathrm{z}}=\sum_0 \overbrace{\mathrm{X}_{\mathrm{o}, \mathrm{t}}}^{\text {Shift }} \cdot \overbrace{\mathrm{N}_{\mathrm{o}, \tau, \mathrm{Z}} / \mathrm{N}_{\tau, \mathrm{z}}}^{\text {Share }}
$$

Where the shares are lagged to a base period τ. Still, the main characteristic is that the shares change at the locality level $z$  and time $t$; in our case $cz$ and $t$. Therefore, my hunch is that the instruments that have shares with a constant time are not delocalized enough to be considered a shift-share instrument. So I will only keep instrument 3 for the analysis. 


## Question 4  {.unnumbered}


::: {.callout-note}
 Compute the estimates of the parameters of the following regression.
$$
\Delta \ell_{c z, t}=\beta \hat{X}_{c z, t}+\gamma_{c z}+\gamma_t+\varepsilon_{c z, t}
$$
where $\gamma_{c z}$ are CZ fixed effects, $\gamma_t$ are year fixed effects, and $\varepsilon_{c z, t}$ is an error term. Pick any $\hat{Z}_{c z, t}^i$ to instrument $\hat{X}_{c z, t}$. Use the OLS and IV estimators with fixed effects. Cluster the standard errors at the CZ-year level. Present a table with four columns: the OLS, the First Stage, the Reduced Form, and the IV regressions. Ensure to include the coefficient of interest, standard error, number of observations, first stage F statistic, and $R^2$.
:::


The strategy is to use [Kolesá'sr](https://cran.r-project.org/web/packages/ShiftShareSE/ShiftShareSE.pdf) package `ShiftShareSE` to avoid the problems anticipated in questions 5-6 regarding the OLS regressor and the correct assumptions to estimate the causal effect.

Yet, I didn't have enough time to review which is the appropriate formula to implement fixed-effects or how to shape the data so it can fit the equation and function.

```{r}
#| eval: false
#| message: false
#| warning: false
#| output: true

library(ShiftShareSE)

formula <- as.formula( "outcome ~ explanatory" )

# Incomplete :( 

# reg_ss( formula , X = z_3, data = model, region_cvar = CZ, , method = "all")


```

 



## Question 5 {.unnumbered}

Discuss in less than 150 words the main concerns to interpret $\beta$ as a causal estimate with the OLS estimator.

According to  [Adão  et al. (2019) ](https://economics.harvard.edu/files/economics/files/ms29827.pdf)   regression residuals are correlated across regions with similar sectoral shares, independently of their geographic location. 


## Question 6 {.unnumbered}

Considering the results of the regressions, discuss in less than 300 words the assumptions for the IV estimator to causally estimate $\beta$ and how you can assess them with the results you obtained.

According to  [Adão  et al. (2019) ](https://economics.harvard.edu/files/economics/files/ms29827.pdf), to estimate the causal effect of regional factors using shift-share designs, several assumptions need to be met, including:

- Exclusion restriction: The regional factor of interest must only affect the outcome variable through its effect on the regional component of the shift-share analysis and not through other channels.

- Monotonicity: The shift-share weights must be monotonic in the regional variable, meaning that regions with higher values of the regional variable must have higher weights than regions with lower values.

- Independence: The shift-share weights must be independent of the regional factor of interest, meaning that regions with higher values of the regional factor cannot have higher weights due to some other unobserved factor.

- Stable treatment: The national factor must be stable over time and not affected by the regional factor of interest.

## Question 7 {.unnumbered}

::: {.callout-note}
Given your last response, discuss in less than 300 words the relevance of the instrument interpreting the different tests' statistics and if the IV strategy works in this context.
:::


I'm out of time :(

This is my first time reading and implementing Shift-Share instruments; I could have delivered a more structured document with more time and preparation. Sorry for the inconvenience while reading the report.




